---
title: "In-Class Exercise 03"
description: |
 R for Geospatial Data Science
author:
  - name: Khant Min Naing
    url: https://www.linkedin.com/in/khantminnaing/
date: 01-12-2024
date-modified: "last-modified"
categories:
  - In-Class Exercise
  - R
  - sf
  - lubridate
  - grab-posisi
format:
  html:  
    #code-fold: true
    code-summary: "Show the code"
execute: 
  eval: true
  echo: true
  warning: false
output:
  distill::distill_article:
    toc: true
    self_contained: false
---

### 1.0 Installing Maptools

```{r}
pacman::p_load(arrow,lubridate,tidyverse,tmap,sf)
```

```{r}
#| eval: False
install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

## 3.0 Importing Datasets to R Environment

In this exercise, we will use the following datasets:

-   `CHILDCARE`, a point feature data providing both location and attribute information of childcare centres. It was downloaded from Data.gov.sg and is in geojson format.

-   `MP14_SUBZONE_WEB_PL`, a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format. This data set was also downloaded from Data.gov.sg.

-   `CostalOutline`, a polygon feature data showing the national boundary of Singapore. It is provided by SLA and is in ESRI shapefile format.

### 3.1 Importing Geospatial Data

In this section, [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package will be used to import the three geospatial data sets mentioned in previous section into R environment.

```{r}
childcare_sf <- st_read("../data/aspatial/child-care-services-geojson.geojson")

mpsz_sf <- st_read(dsn = "../data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

## 4.0 Geospatial Data Wrangling

### 4.1. Assigning Stanadrd Coordinate Systems

```{r}
childcare_sf <- st_transform(childcare_sf, crs = 3414)
mpsz_sf <- st_transform(mpsz_sf, crs = 3414)
```

### 4.2 Creating Coastal Outline

```{r}
sg_sf <- mpsz_sf %>% st_union()
plot(sg_sf)
```
